---
title: "Capstone - Backorders Impact"
author: "Ken Muckey"
date: "1/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Capstone
Overview: Comparing data from 2019 to 2021 to determine gernalized impact of back orders in projects at my workplace.

```{r packages}
install.packages("readxl")
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dplyr")
install.packages("janitor")

library(readxl)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(janitor)
```
## Import

Let's import the raw data from XLSX files...

```{r}
# xlsx files
data_2018 <- read_excel("2018datamgm.xlsx")
data_2021a <-  read_excel("2021datasl.xlsx")
data_2021b <-  read_excel("2021datacosj.xlsx")
data_2022 <-  read_excel("2022datavvs.xlsx")
```

Summaries of each:

```{r}
head(data_2018)
```


```{r}
head(data_2021b)

```

Create new dataframes with selected columns and filtered data:
* Order Date
* Install Date
* Model

```{r}

orders_2018 <- data_2018[c('Order Date','Completed Install Date', 'Toshiba Replacement Model')]
orders_2022 <- data_2022[c('Order Date','Completed Install Date', 'Toshiba Replacement Model')]
```

Create new dataframes and update non-standard columns for one input source.

```{r}
orders_2021a <- data_2021a[c('Order Date','Completed Install Date', 'Toshiba Replacement Model')]
orders_2021b <- data_2021b[c('Order Date','Completed Equipment Install Date', 'Replacement Model' )]
colnames(orders_2021b)[2]<-'Completed Install Date'
colnames(orders_2021b)[3]<-'Toshiba Replacement Model'


```

Clean all column names and adjust one non-standard data source.

```{r}
orders_2018 <- clean_names(orders_2018)
orders_2021a <- clean_names(orders_2021a)
orders_2021b <- clean_names(orders_2021b)
orders_2022 <- clean_names(orders_2022)

orders_2021b$toshiba_replacement_model <- gsub("e-STUDIO","",orders_2021b$toshiba_replacement_model)
head(orders_2021b)

```


The next few sections are filtering dataframes with some unique data conversions per inputs.

```{r filtered_orders_2018}

filtered_orders_2018 <- filter(orders_2018, order_date != "N/A")

filtered_orders_2018$order_date <- excel_numeric_to_date(as.numeric(as.character(filtered_orders_2018$order_date)), date_system = "modern") 

filtered_orders_2018$completed_install_date <- excel_numeric_to_date(as.numeric(as.character(filtered_orders_2018$completed_install_date)), date_system = "modern") 

head(filtered_orders_2018)
```


```{r filtered_orders_2021a}

filtered_orders_2021a <- filter(orders_2021a, order_date != "N/A")

filtered_orders_2021a$order_date <- excel_numeric_to_date(as.numeric(as.character(filtered_orders_2021a$order_date)), date_system = "modern") 

filtered_orders_2021a$completed_install_date <- excel_numeric_to_date(as.numeric(as.character(filtered_orders_2021a$completed_install_date)), date_system = "modern") 


head(filtered_orders_2021a)

```

```{r filtered_orders_2021b}

filtered_orders_2021b <- filter(orders_2021b, order_date != "N/A")

filtered_orders_2021b$order_date <- excel_numeric_to_date(as.numeric(as.character(filtered_orders_2021b$order_date)), date_system = "modern") 

filtered_orders_2021b$completed_install_date <- as.Date(filtered_orders_2021b$completed_install_date)


head(filtered_orders_2021b)

```


```{r filtered_orders_2022}

filtered_orders_2022 <- filter(orders_2022, !is.na(order_date))


filtered_orders_2022$order_date <-as.Date(filtered_orders_2022$order_date)

filtered_orders_2022$completed_install_date <- excel_numeric_to_date(as.numeric(as.character(filtered_orders_2022$completed_install_date)), date_system = "modern") 

head(filtered_orders_2022)

```



Add delta and man to each dataframe for DA later.


```{r}
filtered_orders_2018 <-
  mutate(filtered_orders_2018, delta =  completed_install_date - order_date, mean_delta = mean(delta, na.rm=TRUE))
head (filtered_orders_2018)
```
```{r}

filtered_orders_2021a <-
  mutate(filtered_orders_2021a, delta =  completed_install_date - order_date, mean_delta = mean(delta, na.rm=TRUE))

filtered_orders_2021b <-
  mutate(filtered_orders_2021b, delta =  completed_install_date - order_date, mean_delta = mean(delta, na.rm=TRUE))

filtered_orders_2022 <-
  mutate(filtered_orders_2022, delta =  completed_install_date - order_date, mean_delta = mean(delta, na.rm=TRUE))

head (filtered_orders_2021b)


```

Plot: Display order-to-install number of days per each installation among all data frames.  Colors used to seperate the years.  

```{r}
ggplot() +
  geom_point(data= filtered_orders_2018, aes(x=toshiba_replacement_model, y=delta, color='2018')) +
  geom_point(data= filtered_orders_2021a, aes(x=toshiba_replacement_model, y=delta, color='2021a')) +
  geom_point(data= filtered_orders_2021b, aes(x=toshiba_replacement_model, y=delta, color='2021b')) +
  geom_point(data= filtered_orders_2022, aes(x=toshiba_replacement_model, y=delta, color='2022')) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
```{r}
# ggplot() +
#   geom_bar(data= filtered_orders_2018, aes(x=toshiba_replacement_model, y=delta)) 

delta_summary <- data.frame('by_year' = c('2018','2021a','2021b','2022'),
                            'delta_mean' = c(filtered_orders_2018$mean_delta[1], ... = filtered_orders_2021a$mean_delta[1], filtered_orders_2021b$mean_delta[1], filtered_orders_2022$mean_delta[1])
                            )

ggplot(data=delta_summary, aes(x=by_year, y=delta_mean, fill=by_year)) +
  geom_bar(stat="identity")

#ggplot(data=filtered_orders_2018, aes(x=toshiba_replacement_model, y=delta, fill=toshiba_replacement_model)) +
#  geom_bar(stat="identity")

#ggplot(data=filtered_orders_2018, aes(x=toshiba_replacement_model, y=delta, fill=toshiba_replacement_model)) +
#  geom_bar(stat="identity")


# p <- ggplot(data=df2, aes(x=dose, y=len, fill=supp)) +
# geom_bar(stat="identity", color="black", position=position_dodge())+
#   theme_minimal()
# # Use custom colors
# p + scale_fill_manual(values=c('#999999','#E69F00'))
# # Use brewer color palettes
# p + scale_fill_brewer(palette="Blues")

```
